---
title: "LAK24_FLPS_Gaming"
author: "Kirk Vanacore"
date: "9/7/2023"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Packages
```{r packages}
packages = c(
 # "tidyverse",
  "plyr",
  "rstanarm",
  "ggExtra",
  "xts",
  "lubridate",
  "readxl",
  "data.table",
  "RSQLite",
  "DBI",
  "mice"
  ,"psych",
  "stringr",
  "sjmisc",
  "sjPlot",
  "lme4",
  "campfin",
  "pROC",
  "cutpointr",
  "missForest",
  "corrplot",
  "tidyverse",
  "arm",
  "splines",
  'fastDummies',
  "rstan",
  "arm",
  "doParallel",
  "dplyr",
  "MASS",
  "latex2exp",
  "splines",
  "pROC",
  "xtable"
)
#load install
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only = TRUE)) {
      install.packages(x, dependencies = TRUE)
      library(x, character.only = TRUE)
    }
  } 
) 

require(dplyr)
require(doParallel)
require(rstan)
rm(packages)
cores <- detectCores()
registerDoParallel(cores=10)
```
# Train Measurement Model
This is the process of selecting variables for the measurement model. The final measurement submodel is run similtan

## Train/Test Data 
```{r}
meas_model_data <- read.csv("/Users/kirkvanacore/Documents/WPI_Analyses/FLPS_GamingTheSystem/data/LAK24_Data/meas_model_data.csv")
dim(meas_model_data)
# separate the test and train set 


train_samp <- sample(meas_model_data$ID, size = round(length(meas_model_data$StuID)*.8))
mod1_dat_train <- meas_model_data %>%
  filter(ID %in% train_samp)
mod1_dat_test <- meas_model_data %>%
  filter(!ID %in% train_samp
  ) 

table(mod1_dat_test$gaming_binary)/length(mod1_dat_test$gaming_binary)
table(mod1_dat_train$gaming_binary)/length(mod1_dat_train$gaming_binary)
```

## Measurement Model
```{r}

# check for colinairty
meas_cor<-as.data.frame((cor(as.matrix(mod1_dat_train %>%dplyr::select(
 pre_total_math_score,
  pre_sub_P_score ,
    pre_sub_C_score ,
     pre_sub_F_score ,
     pre_math_completed_num ,
     pre_total_time_on_tasks ,
     pre_MA_total_score ,
    pre_negative_reaction_score ,
    pre_numerical_confindence_score ,
    pre_MSE_total_score ,
    pre_PS_tasks_total_score ,
    pre_PS_part1_score ,
    pre_PS_part2E_score ,
    pre_PS_part2NE_score ,
    pre_PS_completed_num ,
    pre_PS_total_RT_sec_log ,
    pre_PS_part1_RT_sec_log ,
    pre_PS_part2E_RT_sec_log ,
    pre_PS_part2NE_RT_sec_log ,
    Scale_Score5 
   
) ))))
View(meas_cor)
getOption("max.print")

```

```{r}
mm_1null <- stan_glmer(
  gaming_binary ~
    (1 | clip_start_problem) +
    (1 | StuID/TeaIDPre)
  ,
  data = mod1_dat_train,
  chains=5, iter=500, warmup=400, cores = 10,
  family = binomial())

mm_1null_lmer <- glmer(
  gaming_binary ~
    (1 | clip_start_problem) +
    (1 | StuID)
  ,
  data = mod1_dat_train,
 # chains=5, iter=500, warmup=400, cores = 10,
  family = binomial)
summary(mm_1null_lmer)

tab_model(mm_1null_lmer)
mm_null_coef<-mm_1null$stan_summary

# the original process for selecting variables was iterative, the final measrement model is below
mm_1 <- stan_glmer(
  gaming_binary ~
    #pre_total_math_score +
    pre_sub_P_score +
    pre_sub_C_score +
    pre_sub_F_score +
    pre_math_completed_num +
    log(pre_total_time_on_tasks) +
    pre_MA_total_score +
    I(pre_MA_total_score^2) +
    pre_negative_reaction_score +
    pre_numerical_confindence_score +
    pre_MSE_total_score +
    #pre_PS_tasks_total_score +
    pre_PS_part1_score +
    pre_PS_part2E_score +
    pre_PS_part2NE_score +
  #  pre_PS_completed_num +
    pre_PS_total_RT_sec_log +
    pre_PS_part1_RT_sec_log +
    pre_PS_part2E_RT_sec_log +
    pre_PS_part2NE_RT_sec_log +
    Scale_Score5 +
    FEMALE +
    race_ethnicity_Hispanic +
    race_ethnicity_Asian_pacific +
    race_ethnicity_Black +
    IEP +
    EIP +
    ESOL +
    GIFTED +
    inperson +
    na_count +
    (1 | clip_start_problem) +
    (1 | StuID)
  ,
  data = mod1_dat_train,
  chains=5, iter=1000, warmup=900, cores = 10,
  family = binomial())

summary(mm_1)
mm_1$stan_summary

  
#saveRDS(mm_1, "FLPS_GamingTheSystem/models/measurement_mode2mcmc.rda")
mm<- readRDS("/Users/kirkvanacore/Documents/WPI_Analyses/FLPS_GamingTheSystem/models/measurement_models/LAK24_Models/measurement_model2mcmc.rda")

```

```{r}
print(mm2)

mm2_cof <- as.data.frame(mm2$stan_summary)

plot(mm2, "trace", pars = "b[(Intercept) StuID:59]")
plot(mm2, "trace", pars = "b[(Intercept) StuID:59]")
plot(mm2, "trace", pars = "b[(Intercept) StuID:59]")
plot(mm2, "trace", pars = "b[(Intercept) StuID:59]")
```

```{r}
traceplot(
  mm2,
  inc_warmup = T,
  pars = c(
       'b[(Intercept) StuID:59]')
  )

```


```{r}
mm<- readRDS("/Users/kirkvanacore/Documents/WPI_Analyses/FLPS_GamingTheSystem/models/LAK24_Models/measurement_modelmcmc.rda")
summary(mm)
mm$stan_summary
p_mcmc<-(mm$fitted.values)
p_test <-posterior_predict(mm, 
                 newdata = mod1_dat_test,  
                 allow.new.levels = T,
                 type = "response"
)
#p_test
pROC::auc(mm$y, p_mcmc)
pROC::auc(mod1_dat_test$gaming_binary, p_test)

```
#### Variance Explained
This script provides the proportion of variance explained in the model by the variables available to both treatment and control. It serves as a measure of how accurately the model will predict replay behavior in the control group.
```{r prelim_mod_eval} 
# evaluate model - R_sq
student <- mod1_dat_train %>%
dplyr::select(  
   
 # pre_total_math_score,
    pre_sub_P_score,
    pre_sub_C_score,   
    pre_sub_F_score,   
    pre_math_completed_num,  
    pre_total_time_on_tasks,
    pre_MA_total_score,
    pre_negative_reaction_score,
    pre_numerical_confindence_score,
    pre_MSE_total_score,
 #   pre_PS_tasks_total_score,
    pre_PS_part1_score,
    pre_PS_part2E_score,
    pre_PS_part2NE_score,
  #  pre_PS_completed_num,
  #  pre_PS_total_RT_sec_log,
    pre_PS_part1_RT_sec_log,
    pre_PS_part2E_RT_sec_log,
    pre_PS_part2NE_RT_sec_log,
    Scale_Score5,
    FEMALE,
    race_ethnicity_Hispanic,
    race_ethnicity_Asian_pacific,
    race_ethnicity_Black,
    IEP,
    EIP,
    ESOL,
    GIFTED,
    inperson,
    na_count 
   ) %>%
  distinct() 

x = model.matrix(~
 # pre_total_math_score +
    pre_sub_P_score +
    pre_sub_C_score +   
    pre_sub_F_score +   
    pre_math_completed_num +  
    pre_total_time_on_tasks +
    pre_MA_total_score +
    I(pre_MA_total_score^2) +
    pre_negative_reaction_score +
    pre_numerical_confindence_score +
    pre_MSE_total_score +
 #   pre_PS_tasks_total_score +
    pre_PS_part1_score +
    pre_PS_part2E_score +
    pre_PS_part2NE_score +
  #  pre_PS_completed_num +
  #  pre_PS_total_RT_sec_log +
    pre_PS_part1_RT_sec_log +
    pre_PS_part2E_RT_sec_log +
    pre_PS_part2NE_RT_sec_log +
    Scale_Score5 +
    FEMALE +
    race_ethnicity_Hispanic +
    race_ethnicity_Asian_pacific +
    race_ethnicity_Black +
    IEP +
    EIP +
    ESOL +
    GIFTED +
    inperson +
    na_count 
    , data = student)

y_hat = x %*%fixef(mm)

# variance explained by variables used in imputation of alpha
var_alpha<-aphla <-(var(y_hat) + 
    as.data.frame( VarCorr(mm))[1, 4] + 
    as.data.frame( VarCorr(mm))[2, 4])
# variance explained include student
var_aplhaPLUSst<-(var(y_hat) + 
    as.data.frame( VarCorr(mm))[1, 4] + 
    as.data.frame( VarCorr(mm))[2, 4] + 
    as.data.frame( VarCorr(mm))[3, 4] 
    )
# variance explained include student
var_total<-(var(y_hat) + 
    as.data.frame( VarCorr(mm))[1, 4] + 
    as.data.frame( VarCorr(mm))[2, 4] + 
    as.data.frame( VarCorr(mm))[3, 4] + 
    as.data.frame( VarCorr(mm))[4, 4] 
    )

var_alpha/var_aplhaPLUSst
var_alpha/var_total


```

# FLPS Effect of Feedback on Gamers
## Data/Models in STAN
### data
```{r}

saveRDS(stanDat, "/Users/kirkvanacore/Documents/WPI_Analyses/FLPS_GamingTheSystem/data/LAK24_Data/LAK_RQ2_FLPS_gaming_InstVDelay_stan.RData")

str(stanDat)


```


### run models in Secret Sauce
```{r}

FLPS_Mod <- stan('FLPS_GamingTheSystem/code/FLPS_Gaming_2Cond.stan',
            data = stanDat, chains=10, iter=11000, warmup=9000, cores = 10)
saveRDS(RQ2_Mod, "FLPS_GamingTheSystem/models/FLPS_gaming_InstVDelay_Mod4.rds")

print(RQ2_Mod, par = c("a1", "b0", "b1"))

```

### load model
```{r}
FLPS_Mod<-read_rds("/Users/kirkvanacore/Documents/WPI_Analyses/FLPS_GamingTheSystem/models/LAK24_Models/FLPS_gaming_InstVDelay_Mod4.rda")
```

```{r}
summary(FLPS_Mod, par = c("a1", "b0", "b1"))$summary
```


```{r}
a1 = as.data.frame(extract(FLPS_Mod, par = "a1"))
mean((a1$a1) < 0) 

b0 = as.data.frame(extract(FLPS_Mod, par = "b0"))
mean((b0$b0) > 0) 

b1 = as.data.frame(extract(FLPS_Mod, par = "b1"))
mean((b1$b1) < 0)
```


###Model Table
```{r}




#Measurement Model
coef_U<-summary(FLPS_Mod, par = c("betaU" ))$summary
coef_U <- as.data.frame( coef_U)
row.names(coef_U) <- c(
  "Algebraic Procedural Knowledge",
  "Algebraic Conceptual Knowledge",
  "Algebraic Flexibility Knowledge",
  "Algebraic Knowledge Items Complete",
  "Algebraic Knowledge Time (Log)",
  "Math Anxiety",
  "Math Anxiety (Squared)",
  "Math Negitive Reaction",
  "Math Numerical Confindence",
  "Math Self Efficacy",
  "Perceptual Sensitivity Score Part 1",
  "Perceptual Sensitivity Time Part 1 (Log)",
  "Perceptual Sensitivity Score Part 2",
  "Perceptual Sensitivity Time Part 2 (Log)",
  "Perceptual Sensitivity Score Part 3",
  "Perceptual Sensitivity Time Part 4 (Log)",
  "State Test Score",
  "Female",
  "Hispanic",
  "Asian/Pacific Islander",
  "Black",
  "IEP",
  "EIP",
  "ESOL",
  "GIFTED",
  "Inperson Insrutcion",
  "Missing Data"
)

# Probabilities - different from zero
bU<-as.data.frame(extract(FLPS_Mod, par = "betaU"))
head(bU)

prob_gtZero<-function(v) {
  mean(v > 0)
}

prob_ltZero<-function(v) {
  mean(v < 0)
}
mean(bU$betaU.1)
prob_difZero(bU$betaU.1)


mean(bU$betaU.13)
prob_difZero(bU$prob_difZero)


coef_U$`P(>0)` <-unlist(lapply(bU, prob_gtZero))
coef_U$`P(<0)` <-unlist(lapply(bU, prob_ltZero))

colnames(coef)

coef_U<-coef_U  %>%
  mutate("Predictors"=rownames(.),
         " " = " ") %>%
  dplyr::select(
    "Predictors",     "Estimate" = mean,
         "SD" = sd,
         "P(>0)",
         "P(<0)",
          " " 
         
         )




#Measurement Model
coef_Y<-summary(FLPS_Mod, par = c("betaY" ))$summary
coef_Y <- as.data.frame( coef_Y)
row.names(coef_Y) <- c(
 "Algebraic Procedural Knowledge",
  "Algebraic Conceptual Knowledge",
  "Algebraic Flexibility Knowledge",
  "Algebraic Knowledge Items Complete",
  "Algebraic Knowledge Time (Log)",
  "Math Anxiety",
  "Math Anxiety (Squared)",
  "Math Negitive Reaction",
  "Math Numerical Confindence",
  "Math Self Efficacy",
  "Perceptual Sensitivity Score Part 1",
  "Perceptual Sensitivity Time Part 1 (Log)",
  "Perceptual Sensitivity Score Part 2",
  "Perceptual Sensitivity Time Part 2 (Log)",
  "Perceptual Sensitivity Score Part 3",
  "Perceptual Sensitivity Time Part 4 (Log)",
  "State Test Score",
  "Female",
  "Hispanic",
  "Asian/Pacific Islander",
  "Black",
  "IEP",
  "EIP",
  "ESOL",
  "GIFTED",
  "Inperson Insrutcion",
  "Missing Data"
)


# Probabilities - different from zero
bY<-as.data.frame(extract(RQ2_Mod, par = "betaY"))



coef_Y$`P(>0)` <-unlist(lapply(bY, prob_gtZero))
coef_Y$`P(<0)` <-unlist(lapply(bY, prob_ltZero))


coef_Y<-coef_Y  %>%
  mutate("Predictors"=rownames(.)) %>%
  dplyr::select(
    "Predictors",     "Estimate" = mean,
         "SD" = sd,
         "P(>0)",
         "P(<0)"
         
         )

cbind(coef_U, coef_Y)

# effects/interactions
alpha_beta<-summary(FLPS_Mod, par = c("a1", "b0", "b1"))$summary
alpha_beta <- as.data.frame( alpha_beta)

ab<-as.data.frame(extract(FLPS_Mod, par =c("a1", "b0", "b1")))

alpha_beta$`P(>0)` =unlist(lapply(ab, prob_gtZero))
alpha_beta$`P(<0)` =unlist(lapply(ab, prob_ltZero))


alpha_beta <- alpha_beta %>%
  mutate("Predictors"=rownames(.)) %>%
  dplyr::select(
    "Predictors",     
    "Estimate" = mean,
         "SD" = sd,
         "P(>0)",
         "P(<0)"
    )

table1<-right_join(coef_U, 
(rbind(alpha_beta, coef_Y)),
by = "Predictors") %>%
  mutate(across(where(is.numeric), round, 3))
colnames(table1) <- c("Predictors", "Estimate", "SD", "P(>0)", "P(<0)", "", "Estimate","SD","P(>0)","P(<0)")

View(table1)
write_csv(table1, "Table1.csv")
print(xtable(table1,
             label = "Fully Latent Principal Statsification Model Parameter Estimates",
             align = "lcccccccccc"),
      include.rownames=FALSE)
print(xtable(table1,
             label = "Fully Latent Principal Statsification Model Parameter Estimates",
             align = "llccccccccc"),
      include.rownames=FALSE)
```


#### usage model


```{r}
studEff<-summary(RQ2_Mod, 
  pars = c(
    'studEff'),
  probs = c(0.025, 0.975))

studEff <- as.data.frame(studEff$summary)

stats::var(studEff$mean)
mean(studEff$mean)
quantile(studEff$mean)

studEff$Q <- ifelse(studEff$mean < quantile(studEff$mean)[2], 1,
                    ifelse(studEff$mean < quantile(studEff$mean)[3], 2,
                    ifelse(studEff$mean < quantile(studEff$mean)[4], 3,
                    4
                           )))
View(studEff %>% group_by(Q) %>% summarise(alpha = round(mean(mean), 2),
                                      tau = round(mean(b0$b0)+alpha*(mean(b1$b1)), 2)))
     
     
mean(, group = studEff$Q)

mean(studEff$summary[,1] >=1)
max(studEff$summary[,1])

```

```{r}
stud_mu<-summary(RQ2_Mod, 
  pars = c(
    'mu_prob'),
  probs = c(0.025, 0.975))

```


```{r}
hist(exp(studEff$summary[,1])/(1+exp(studEff$summary[,1])))


studEff_prb <-
  cbind(as.data.frame(exp(studEff$summary[, 1]) / (1 + exp(
    studEff$summary[, 1]
  ))), (studEff$summary[, 1]), stanDat$Z, stanDat$Y)

colnames(studEff_prb)[1] <- "alpha_prob"
colnames(studEff_prb)[2] <- "alpha_logit"
colnames(studEff_prb)[3] <- "Z"
colnames(studEff_prb)[4] <- "Y"

describe(studEff_prb$alpha_logit)
describe(studEff_prb$alpha_logit)
mean(studEff_prb$alpha_logit >= 1)
quantile(studEff_prb$alpha_logit)



wilcox.test(studEff_prb$alpha_logit ~  as.factor(studEff_prb$Z))
describeBy(studEff_prb$alpha_logit, as.factor(studEff_prb$Z))





library(latex2exp)

ggplot(studEff_prb[studEff_prb$Z==1, ], aes(x =(`alpha_logit`))) +
  geom_density( alpha = .8, fill = "gray") +
  xlab(TeX("$\\alpha_i$")) +
  theme_classic()

ggplot(studEff_prb, aes(x =`alpha_logit`, fill = as.factor(Z))) +
  geom_histogram( alpha = .5, position = "identity") +
  xlab(TeX("$\\alpha_i$")) +
  theme_classic()


```

### Pronesity to replay 


### traceplots


```{r}
traceplot(
  RQ2_Mod,
  inc_warmup = T,
  pars = c(
       'a1',
    'b0',
    'b1')
  )
traceplot(
  RQ2_Mod,
  inc_warmup = T,
  pars = c(
       'studEff[1]',
       'studEff[2]',
       'studEff[3]',
       'studEff[4]',
       'studEff[5]',
       'studEff[6]',
       'studEff[7]',
       'studEff[8]'
       )
  )
```



